DATA 624 Project 1 Part A - ATM

### 2020-03-13 DATA624/Project 1-A ATM.Rmd
More work on Project 1

```{r}
train_test_split <- function(atm_df, atm_number, n_in_test = 31, return_ts=TRUE){
  ### Splits the data into training and test sets
  # Args:
  #     atm_df (data.frame): The ATM data
  #     atm_number (int): Which ATM do you want the time series for?
  #     n_in_test (int): The number of observations in the test set taken from the end (optional-31 default)
  #     return_ts (logical): Should the function return ts objects (optional-TRUE default)
  # Returns:
  #     train_test_ts (list): Two ts objects (train and test).

@@

  # Split the data frame into training and test sets
  test <- tail(df, n_in_test)
  train <- head(df, -n_in_test)

  # Tranform into ts objects
  if(return_ts){
    train_start <- (min(train$DATE))
    train <- ts(train$Cash, start = decimal_date(as.Date(train_start)), frequency = 365)
    test_start <- (min(test$DATE))
    test <- ts(test$Cash, start = decimal_date(as.Date(test_start)), frequency = 365)
  }

  # Return the data to the user
  train_test <- list(train = train, test = test)
  return(train_test)
}

atm_1_ts <- train_test_split(adjusted_atm, 1, 31)
atm_2_ts <- train_test_split(adjusted_atm, 2, 31)
atm_3_ts <- train_test_split(adjusted_atm, 3, 31)
atm_4_ts <- train_test_split(adjusted_atm, 4, 31)

atm_1_df <- train_test_split(adjusted_atm, 1, 31, return_ts = FALSE)
atm_2_df <- train_test_split(adjusted_atm, 2, 31, return_ts = FALSE)
atm_3_df <- train_test_split(adjusted_atm, 3, 31, return_ts = FALSE)
atm_4_df <- train_test_split(adjusted_atm, 4, 31, return_ts = FALSE)
```

@@

I will start by looking at the training data and the decomposition.  I cannot use the `decompose()` or `stl()` functions to decompose it because I don't have enough data.  Facebook Prophet can decompose it using its Basian process.

@@

>>>>

```{r}

forecast_window <- 31

df_to_ts <- function(df){
  ### Converts ATM data frames into ts objects
  # Args:
  #     df (data.frame): The ATM data frame 
  # Returns:
  #     ts (Time-Series): An ATM time series object
  ts(df$Cash, start = decimal_date(as.Date(min(df$DATE))), frequency = 365)
}

train_test_split <- function(atm_df, atm_number, n_in_test, return_ts=TRUE){
  ### Splits the data into training and test sets
  # Args:
  #     atm_df (data.frame): The ATM data
  #     atm_number (int): Which ATM do you want the time series for?
  #     n_in_test (int): The number of observations in the test set taken from the end (optional-31 default)
  #     return_ts (logical): Should the function return ts objects (optional-TRUE default)
  # Returns:
  #     train_eval_test_ts (list): Data split into sets (train, eval and test).

@@

  # Split the data frame into training, eval and test sets
  test <- tail(df, n_in_test)
  train <- head(df, -n_in_test)
  eval_ts <- tail(train, n_in_test)
  short_train <- head(df, -(2 * n_in_test))

  # Tranform into ts objects
  if(return_ts){
    train <- df_to_ts(train)
    test <- df_to_ts(test)
    eval_ts <- df_to_ts(eval_ts)
    short_train <- df_to_ts(short_train)
  }

  # Return the data to the user
  train_eval_test_ts <- list(train = train, test = test, evaluation = eval_ts, short_train = short_train)

  return(train_eval_test_ts)
}

# Get time series objects
atm_1_ts <- train_test_split(adjusted_atm, 1, forecast_window)
atm_2_ts <- train_test_split(adjusted_atm, 2, forecast_window)
atm_3_ts <- train_test_split(adjusted_atm, 3, forecast_window)
atm_4_ts <- train_test_split(adjusted_atm, 4, forecast_window)
# Get data frame objects
atm_1_df <- train_test_split(adjusted_atm, 1, forecast_window, return_ts = FALSE)
atm_2_df <- train_test_split(adjusted_atm, 2, forecast_window, return_ts = FALSE)
atm_3_df <- train_test_split(adjusted_atm, 3, forecast_window, return_ts = FALSE)
atm_4_df <- train_test_split(adjusted_atm, 4, forecast_window, return_ts = FALSE)
```

@@

I will start by looking at the training data and the decomposition.  I cannot use the `decompose()` or `stl()` functions to decompose it because I don't have enough data.  Facebook Prophet can decompose it using its Bayesian estimation process.

@@

```{r}

atm_1_df$train %>%
  mutate(`Day of Week` = recode(as.factor(wday(DATE)), "1" = "Sunday", "2" = "Monday", "3" = "Tuesday", "4"="Wednesday", "5"="Thursday", "6"="Friday","7"="Saturday")) %>%
  ggplot(aes(x=DATE, y=Cash, color=`Day of Week`)) +
  geom_point() +
  #geom_line() +
  facet_wrap(.~`Day of Week`) +
  theme(legend.position = "None", axis.title = element_blank())
```

Saturday looks fairly consistent.  Other days (Tuesday and Thursday) show recent changes in their trend.

```{r}
atm_1_holt <- holt(atm_1_ts$train, h=forecast_window)
autoplot(atm_1_holt) + autolayer(atm_1_ts$test)
```

```{r}
plot(atm_1_prophet$model, atm_1_prophet$forecast)
```

```{r}
atm_1_mean <- mean(atm_1_df$train)
atm_1_last_observation <- tail(atm_1_df$train, 1)$Cash
atm_1_mean <- mean(atm_1_df$train$Cash)
atm_1_results <- 
  atm_1_prophet$forecast %>%
  mutate(DATE = as.Date(ds), prophet_yhat = yhat) %>%
  select(DATE, prophet_yhat) %>%
  mutate(naive_yhat = atm_1_last_observation, 
         mean_yhat = atm_1_mean) %>%
  merge(atm_1_df$evaluation) %>%
  rename(actual = Cash)
atm_1_results$holt <- tail(atm_1_holt$fitted, forecast_window)
```
